Parallel air palletization and tour planning for simultaneous pickup and delivery

### 2023-02-04 latex/comp_or.tex
mpShims still with values "better" than optimal

The final experiments are performed on a 64-bit, 128GiB, 3.6GHz, 48-core processor, 2 threads per core, Intel Xeon, with {\it Linux CentOS 8} as the operational system and {\it Python 3.8.2} as the programming language.

For parallelization, we used the {\it Python Multi-processing} package, which implements {\it process-based parallelism} or {\it parallel shared memory algorithm}. According to \cite{Multi-processing}, {\it Multi-processing} is a package that supports spawning processes using an API similar to the {\it Python Threading} package.

@@

By Amdahl's law, there is an optimal number of parallel executions for each problem in each environment. While working at IBM in 1967, Gene Amdahl developed the foundation for what became known as Amdahl's Law or Amdahl's Argument. Essentially, the law states that while a process can be decomposed into steps that may then be run in parallel, the time taken for the whole process will be significantly limited by the steps that remain serialized. So, we decided to run some tests to find the best number of parallel processes for each of this work's operational scenarios.

For each red bullet closest to the upper left corner, we call it {\it Utopia}, an impossible result with zero time duration and a possible best score. The blue bullets are experiments performed with the number of parallel processes indicated. The shortest distance from {\it Utopia} may reveal the most efficient parallel processes for the scenario (see Table \ref{tab:scenarios}).

\begin{figure}[H]
	\begin{center}
		\begin{tikzpicture}[scale=0.8, samples=100]
			\begin{axis}[
				myplotstyle,
				xlabel={\bf mpShims computation times (s)},
				ylabel={\bf Score},
				]
				\addplot+[
				only marks,	
				nodes near coords = {\numProcs{}},
				visualization depends on = {value \thisrow{np} \as \numProcs},						
				]
				table[x=time, y=score, col sep=comma] {./csv/Shims_p1data20.csv};
			\end{axis}
			\fill [red] (canvas cs:x=0.1cm,y=5.6cm) circle (2pt);
		\end{tikzpicture}%
		~%
		%
		\begin{tikzpicture}[scale=0.8, samples=100]
		\begin{axis}[
			myplotstyle,
			xlabel={\bf mpACO computation times (s)},
			ylabel={\bf Score},
			]
			\addplot+[
			only marks,	
			nodes near coords = {\numProcs{}},
			visualization depends on = {value \thisrow{np} \as \numProcs},						
			]
			table[x=time, y=score, col sep=comma] {./csv/ACO_p1data20.csv};
		\end{axis}
		\fill [red] (canvas cs:x=0.1cm,y=5.6cm) circle (2pt);
	\end{tikzpicture}

	\end{center}
	\caption{ Scenario 1 performances with $1.2 \times volume$. }
	\label{fig:scenario1data20}
\end{figure}

Observing Figure \ref{fig:scenario1data20}, it may be noticed that the best number of processes for {\it Scenario 1} solved with {\it mpShims} (the closest to the {\it Utopia}) is \textbf{6}. Solved with {\it mpACO}, the best number is \textbf{8}. Table \ref{tab:numProcs} presents all methods, volumes percentages and scenarios numbers of best parallel processes to be run to solve all tours.

\begin{figure}[H]
	\begin{center}
		\begin{tikzpicture}[scale=0.8, samples=100]
			\begin{axis}[
				myplotstyle,
				xlabel={\bf mpShims computation times (s)},
				ylabel={\bf Score},
				]
				\addplot+[
				only marks,	
				nodes near coords = {\numProcs{}},
				visualization depends on = {value \thisrow{np} \as \numProcs},						
				]
				table[x=time, y=score, col sep=comma] {./csv/Shims_p1data50.csv};
			\end{axis}
			\fill [red] (canvas cs:x=0.1cm,y=5.6cm) circle (2pt);
		\end{tikzpicture}%
		~%
		%
		\begin{tikzpicture}[scale=0.8, samples=100]
			\begin{axis}[
				myplotstyle,
				xlabel={\bf mpACO computation times (s)},
				ylabel={\bf Score},
				]
				\addplot+[
				only marks,	
				nodes near coords = {\numProcs{}},
				visualization depends on = {value \thisrow{np} \as \numProcs},						
				]
				table[x=time, y=score, col sep=comma] {./csv/ACO_p1data50.csv};
			\end{axis}
			\fill [red] (canvas cs:x=0.1cm,y=5.6cm) circle (2pt);
		\end{tikzpicture}

	\end{center}
	\caption{ Scenario 1 performances with $1.5 \times volume$. }
	\label{fig:scenario1data50}
\end{figure}

To build Table \ref{tab:numProcs}, it was necessary to determine the number of parallel processes that best solve each node. For this we created a procedure as in Algorithm \ref{alg:bestnum}.

\begin{algorithm}[H]
	\caption{Get the best number of parallel processes}  \label{alg:bestnum}
	\begin{algorithmic}[1]

		\Procedure{$getBestNPP$}{$scores,\ runtimes,\ procs$}

		\State $maxS   \gets max(scores)$
		\State $minS   \gets min(scores)$
		\State $maxR \gets max(runtimes)$
		\State $minR \gets min(runtimes)$

		\State $bestNPP \gets 0$, best number of parallel processes
		\State $minD \gets 999999999.9$

		\For{$i \in |scores|$}
			\State $y = ( maxS - scores[i] )\ /\ (maxS   - minS  )$ \Comment{the Y axis leg}
			\State $x = ( runtimes[i] - minR )\ /\ (maxR - minR)$ \Comment{the X axis leg}
			\State $h = \sqrt{x^2 + y^2}$  \Comment{distance from {\it Utopia}}
			\If{$minD$ > $h$}
				\State $minD \gets h$
				\State $bestNPP \gets procs[i]$
			\EndIf

		\Return $bestNPP$

\begin{table}[H]
\centering
\caption{$bestNPP$ found for each scenario}  \label{tab:numProcs}
	\footnotesize
%\scriptsize
\begin{tabular}{rccccccc}
\toprule
                               &               &\multicolumn{6}{c}{\bf Scenarios} \\
                               \cmidrule{3-8}
{$method$}                     &{\it volume}&{\bf 1}&{\bf 2}&{\bf 3}&{\bf 4}&{\bf 5}&{\bf 6} \\
\toprule

\multirow{3}{*}{\it {mpShims}} &$1.2$& 6       & 4     & 4     & 4     & 2     & 2    \\ 
                               &$1.5$& 4       & 4     & 4     & 4     & 4     & 4    \\
                               &$2.0$& 4       & 4     & 4     & 2     & 2     & 4    \\
\midrule[.1pt]

\multirow{3}{*}{\it {mpACO}}   &$1.2$& 10      & 8     & 10    & 6     & 10    & 10   \\ 
                               &$1.5$& 10      & 6     & 12    & 8     & 10    & 8    \\
                               &$2.0$& 10      & 8     & 2     & 10    & 6     & 6    \\                               
\bottomrule		
\end{tabular}

\normalsize

\end{table}

Table \ref{tab:numProcs} presents the best numbers of parallel processes for each method, scenario, and volume surplus. As all numbers are less than or equal to 12, any hand-held computer with 8 cores can yield efficient solutions, as normally each core may be capable of 2 threads.

To find these values we submitted to test 1, 2, 4, 6, 8, 10, 12, 14, 16 parallel processes.

@@

\multirow{10}{*}{$1.2$} &\multirow{2}{*}{ $Shims$}  & $f$                &       &       &        &       &       &       &   \\%
&                                                   & {\bf run-time (s)} & < 1   &       &        &       &       &       &   \\%
\cmidrule{2-10}
&\multirow{2}{*}{ $mpShims$}                        & $f$                &       &       &        &       &       &       &   \\%
&                                                   & {\bf run-time (s)} & < 1   &       &        &       &       &       &   \\%

\cmidrule{2-10}
&\multirow{2}{*}{ $ACO$}                            & $f$                & 2.85  &       &        &       &       &       &   \\%
&                                                   & {\bf run-time (s)} & 15    &       &        &       &       &       &   \\%
\cmidrule{2-10}		                       
&\multirow{2}{*}{ $mpACO$}                          & $f$                & 3.63  &       &        &       &       &       &   \\%
&                                                   & {\bf run-time (s)} & 11    &       &        &       &       &       &   \\%

\cmidrule{2-10}		                       
&\multirow{2}{*}{ $Gurobi$}                         & $f$                & 6.35  &       &        &       &       &       &   \\%
&                                                   & {\bf run-time (s)} & 2     &       &        &       &       &       &   \\%

\midrule
\multirow{10}{*}{$1.5$} &\multirow{2}{*}{ $Shims$}  & $f$                &       &       &        &       &       &       &   \\%
&                                                   & {\bf run-time (s)} & < 1   &       &        &       &       &       &   \\%
\cmidrule{2-10}
&\multirow{2}{*}{ $mpShims$}                        & $f$                &       &       &        &       &       &       &   \\%
&                                                   & {\bf run-time (s)} & < 1   &       &        &       &       &       &   \\%

\cmidrule{2-10}
&\multirow{2}{*}{ $ACO$}                            & $f$                & 2.85  &       &        &       &       &       &   \\%
&                                                   & {\bf run-time (s)} & 15    &       &        &       &       &       &   \\%
\cmidrule{2-10}		                       
&\multirow{2}{*}{ $mpACO$}                          & $f$                & 3.63  &       &        &       &       &       &   \\%
&                                                   & {\bf run-time (s)} & 11    &       &        &       &       &       &   \\%

\cmidrule{2-10}		                       
&\multirow{2}{*}{ $Gurobi$}                         & $f$                & 6.35  &       &        &       &       &       &   \\%
&                                                   & {\bf run-time (s)} & 2     &       &        &       &       &       &   \\%

\midrule
\multirow{10}{*}{$2.0$} &\multirow{2}{*}{ $Shims$}  & $f$                &       &       &        &       &       &       &   \\%
&                                                   & {\bf run-time (s)} & < 1   &       &        &       &       &       &   \\%
\cmidrule{2-10}
&\multirow{2}{*}{ $mpShims$}                        & $f$                &       &       &        &       &       &       &   \\%
&                                                   & {\bf run-time (s)} & < 1   &       &        &       &       &       &   \\%

\cmidrule{2-10}
&\multirow{2}{*}{ $ACO$}                            & $f$                & 2.85  &       &        &       &       &       &   \\%
&                                                   & {\bf run-time (s)} & 15    &       &        &       &       &       &   \\%
\cmidrule{2-10}		                       
&\multirow{2}{*}{ $mpACO$}                          & $f$                & 3.63  &       &        &       &       &       &   \\%
&                                                   & {\bf run-time (s)} & 11    &       &        &       &       &       &   \\%

\cmidrule{2-10}		                       
&\multirow{2}{*}{ $Gurobi$}                         & $f$                & 6.35  &       &        &       &       &       &   \\%
&                                                   & {\bf run-time (s)} & 2     &       &        &       &       &       &   \\%

@@

\end{tabular}
%\textsuperscript{The numbers 0.7s and 60s as superscript on the methods names represent the time limit per node.}	
\normalsize
\end{table}

>>>>

The final experiments are performed on a 64-bit, 16GiB, 3.6GHz, 8-core AMD® Fx-8320e processor, 2 threads per core, with {\it Linux Ubuntu 20} as the operational system and {\it Python 3.8.2} as the programming language.

For parallelization, we used the {\it Python Multi-processing} package, which implements {\it process-based parallelism} or {\it parallel shared memory algorithm}. According to \cite{multiprocessing}, {\it Multi-processing} is a package that supports spawning processes using an API similar to the {\it Python Threading} package.

@@

By Amdahl's law, there is an optimal number of parallel executions for each problem in each environment. While working at IBM in 1967, Gene Amdahl developed the foundation for what became known as Amdahl's Law or Amdahl's Argument. Essentially, the law states that while a process can be decomposed into steps that may then be run in parallel, the time taken for the whole process will be significantly limited by the steps that remain serialized.

%
%\begin{figure}[H]
%	\begin{center}
%		\begin{tikzpicture}[scale=0.8, samples=100]
%			\begin{axis}[
%				myplotstyle,
%				xlabel={\bf mpShims computation times (s)},
%				ylabel={\bf Score},
%				]
%				\addplot+[
%				only marks,	
%				nodes near coords = {\numProcs{}},
%				visualization depends on = {value \thisrow{np} \as \numProcs},						
%				]
%				table[x=time, y=score, col sep=comma] {./csv/Shims_p1data20.csv};
%			\end{axis}
%			\fill [red] (canvas cs:x=0.1cm,y=5.6cm) circle (2pt);
%		\end{tikzpicture}%
%		~%
%		%
%		\begin{tikzpicture}[scale=0.8, samples=100]
%		\begin{axis}[
%			myplotstyle,
%			xlabel={\bf mpACO computation times (s)},
%			ylabel={\bf Score},
%			]
%			\addplot+[
%			only marks,	
%			nodes near coords = {\numProcs{}},
%			visualization depends on = {value \thisrow{np} \as \numProcs},						
%			]
%			table[x=time, y=score, col sep=comma] {./csv/ACO_p1data20.csv};
%		\end{axis}
%		\fill [red] (canvas cs:x=0.1cm,y=5.6cm) circle (2pt);
%	\end{tikzpicture}
%		
%	\end{center}
%	\caption{ Scenario 1 performances with $1.2 \times volume$. }
%	\label{fig:scenario1data20}
%\end{figure}

@@

\multirow{6}{*}{$1.2$} &\multirow{2}{*}{ Gurobi}& $f$                & 9.75       &  6.34    &  7.61  &   7.43    &       &       &   \\%
&                                               & {\bf run-time (s)} &  8         &   17     &  30    &  114     &       &       &   \\%

\cmidrule{2-10}		                       
&\multirow{2}{*}{ $Shims$}                      & $f$                & 7.00       &  4.20    &  4.12  &  3.54     &       &       &   \\%
&                                               & {\bf run-time (s)} &  < 1       &   3      &    5   &   19    &       &       &   \\%

\cmidrule{2-10}		                       
&\multirow{2}{*}{ $mpShims$}                    & $f$                & {\bf 9.85} &  7.52    &  8.33  &  7.82     &       &       &   \\%
&                                               & {\bf run-time (s)} &  < 1       &  < 1     &    2   &  8     &       &       &   \\%

\midrule

@@

\end{tabular}
%\textsuperscript{$Gurobi^*$ was run with a time limit of 60s per node.}	
\normalsize
\end{table}

$Gurobi^*$ was run with a time limit of 60s per node. The remaining methods were run with a time limit of 1.6s per node, the worst time {\it Shims} took to solve a node.
